---
title: "Data Exploration Hunter Elliott"
author: "Hunter"
format: docx
editor: visual
---

## 

## Liberties

```{r}
library(tidyverse)
library(fixest)
library(rio)
library(lubridate)
library(readr)
library(dplyr)

```

Importing Google Trends Data

Create a vector of file names for all 'trends_up_to\_' files with full paths

```{r}
# Create a vector of filenames for all 'trends_up_to_' files with full paths
file_names <- list.files(pattern = "trends_up_to_", full.names = TRUE)

# Import all files and combine them into a single data frame
combined_data <- import_list(file_names, rbind = TRUE, fill = TRUE)
```

Trim the date string to the correct format

```{r}
combined_data <- combined_data %>%
  mutate(
    date_string = str_sub(monthorweek, end = 10),
    date = ymd(date_string),
    # Floor the date to the first of the month
    first_day_of_month = floor_date(date, unit = "month")
  )
```

Process the combined_data to extract dates and calculate the z-score

```{r}
combined_data <- combined_data %>%
  group_by(schname, keyword) %>%
  mutate(z_score = (index - mean(index, na.rm = TRUE)) / sd(index, na.rm = TRUE)) %>%
  ungroup()
```

Aggregate to month level by taking the average of z-scores

```{r}
combined_data <- combined_data %>%
  group_by(schname, first_day_of_month) %>%
  summarize(average_z_score = mean(z_score, na.rm = TRUE), .groups = 'drop')
```

Import the College Scorecard data

Merge with Scorecard and ID Name Link Data

```{r}
scorecard_data <- import("Most+Recent+Cohorts+(Scorecard+Elements).csv")

# Import the ID name link data
id_name_link <- import("id_name_link.csv")

```

Filter out non-unique school names

Join with_id with scorecard_data on unitid and opeid

```{r}
id_name_link_unique <- id_name_link %>%
  group_by(schname) %>%
  mutate(n = n()) %>%
  filter(n == 1) %>%
  ungroup()

with_id <- inner_join(combined_data, id_name_link_unique, by = "schname")

final_data <- inner_join(with_id, scorecard_data, by = c("unitid" = "UNITID", "opeid" = "OPEID"))
```

Filter **`final_data`** for institutions predominantly granting bachelor's degrees

```{r}
final_data <- final_data %>%
  filter(PREDDEG == 3)
```

Remove rows where earnings data is suppressed or missing

Handle any NAs introduced by coercion and create a high_earning indicator

```{r}
final_data <- final_data %>%
  mutate(md_earn_wne_p10_REPORTED_EARNINGS = as.character(`md_earn_wne_p10-REPORTED-EARNINGS`)) %>%
  # Replace 'PrivacySuppressed' and any other non-numeric strings with NA
  mutate(md_earn_wne_p10_REPORTED_EARNINGS = na_if(md_earn_wne_p10_REPORTED_EARNINGS, "PrivacySuppressed")) %>%
  mutate(md_earn_wne_p10_REPORTED_EARNINGS = ifelse(md_earn_wne_p10_REPORTED_EARNINGS %in% c("", "NULL"), NA, md_earn_wne_p10_REPORTED_EARNINGS)) %>%
  # Now safely convert to numeric
  mutate(md_earn_wne_p10_REPORTED_EARNINGS = as.numeric(md_earn_wne_p10_REPORTED_EARNINGS))
```

create a high_earning indicator

```{r}
final_data <- final_data %>%
  mutate(high_earning = ifelse(!is.na(md_earn_wne_p10_REPORTED_EARNINGS) & md_earn_wne_p10_REPORTED_EARNINGS >= 75000, 1, 0))


```

Select only the necessary variables for the regression analysis

```{r}
analysis_data <- final_data %>%
  select(schname, first_day_of_month, average_z_score, PREDDEG, high_earning, md_earn_wne_p10_REPORTED_EARNINGS)

```

```{r}
analysis_data <- analysis_data %>%
  mutate(post_scorecard = ifelse(first_day_of_month >= as.Date("2015-09-01"), 1, 0))

```

```{r}
# Run the regression model
model <- lm(average_z_score ~ high_earning * post_scorecard, data = analysis_data)
summary(model)

```

Graph to visualize

```{r}
ggplot(analysis_data, aes(x = first_day_of_month, y = average_z_score, group = group, color = group)) +
  geom_line() +  # Plot lines for each group
  labs(title = "Google Search Interest Over Time",
       x = "Date",
       y = "Average Z-Score",
       color = "Group") +
  theme_minimal() +  # Use a minimal theme
  theme(legend.position = "bottom")  # Move the legend to the bottom

```

**Analysis of the Impact of the College Scorecard Release on Google Search Interest**

**Introduction**

In September 2015, the College Scorecard was released, aiming to provide crucial information about colleges, including the earnings of their graduates. This research investigates whether the release of the College Scorecard influenced student interest towards colleges that predominantly grant bachelor's degrees, particularly focusing on the distinction between high-earning and low-earning institutions as reflected through Google search trends.

**Data Preparation and Preliminary Analysis**

The initial step involved filtering the dataset to focus on institutions predominantly granting bachelor's degrees. A significant preprocessing task was to address entries marked as "PrivacySuppressed" and ensure all data pertaining to earnings were numeric md_earn_wne_p10_REPORTED_EARNINGS. The definition of "high-earning" colleges was set at institutions with median earnings above \$75,000, based on findings from the College Scorecard data dictionary "Percent of high-income (above \$75,000 in nominal family income)" variable HI_INC_UNKN_ORIG_YR3_RT and corroborated by median household income data from external sources <https://www.nerdwallet.com/article/finance/median-household-income#:~:text=The%20national%20median%20household%20income,What%20is%20the%20minimum%20wage%3F>. A binary variable was then created to categorize colleges into high-earning (1) and low-earning (0) groups.

**Regression Analysis**

The regression model was designed as follows: model \<- lm(average_z_score \~ high_earning \* post_scorecard, data = analysis_data)

-   **Dependent Variable**: The **`average_z_score`**, representing standardized Google search interest for colleges.

-   **Independent Variables**:

    -   **`high_earning`**: A binary indicator for high-earning colleges.

    -   **`post_scorecard`**: A binary indicator for the period after the College Scorecard's release.

-   **Interaction Term**: **`high_earning * post_scorecard`** to examine differential impacts post-release.

**Results**

-   **Baseline Interest**: The intercept (0.054815) indicates the baseline search interest for low-earning colleges before the Scorecard's release.

-   **Impact on High-Earning Colleges**: Contrary to expectations, high-earning colleges initially showed a slightly lower search interest compared to low-earning colleges, though this difference was not statistically significant (p = 0.448).

-   **Post-Scorecard Release**: There was a notable decline in search interest for low-earning colleges following the Scorecard's release (p \< 2e-16).

-   **Differential Effect**: The interaction term (p = 0.171) suggests a positive, albeit not statistically significant, increase in search interest for high-earning colleges relative to low-earning ones post-release.

**Conclusion**

The introduction of the College Scorecard **decreased** search activity on Google Trends for colleges with high-earning graduates by **0.059900 units** relative to what it did for colleges with low-earning graduates, with a standard error of **0.043712**. This result comes from the **high_earning:post_scorecard interaction** coefficient in my regression.
